Supplementary: A temporal network analysis of drug co-prescription around antidepressants and anxiolytics uses in the Netherlands from 2018 to 2022

### 2025-05-07 docs/abstract/_abstract-short.qmd
Focus the manuscript on co-prescription instead of polypharmacy

**Background:**
Drug prescription networks (DPNs) model the temporal dynamics of medication co-prescription within a population. Understanding these networks can provide insights into polypharmacy and prescribing behaviors.

**Objectives:**
This study assesses the structural characteristics of temporal DPNs derived from daily co-prescriptions of antidepressants, anxiolytics, and other therapeutic drug classes. By analyzing these networks using eigenvector centrality, we identify influential medications and prescribing patterns.

**Methods:**
We utilized the IADB.nl database, including prescriptions from 128 Dutch pharmacies (2018–2022). A closed cohort of patients prescribed antidepressants/anxiolytics was extracted. Medications were classified using the Anatomical Therapeutic Chemical (ATC) system into 24 therapeutic classes. Time-varying DPNs were constructed as undirected graphs using symmetric daily dose-adjusted co-prescriptions. Eigenvector centrality ($c_e$) quantified relative nodal importance. Weekly-aggregated data included prescriptions ($n_c$), number of patients, claim-to-patient ratio, and eigenvector centrality. Singular spectrum analysis decomposed trends.

**Results:**
Antidepressants (ATC `N06A`, $c_e$: `r N06A$ce`, $n_c$: `r N06A$nc`) and anxiolytics (`N05B`, $c_e$: `r N05B$ce`, $n_c$: `r N05B$nc`) had high eigenvector centrality, indicating frequent co-prescription. Additional high-centrality groups included medications for alimentary and metabolism (`A01-A16`), blood (`B01-B06`), cardiovascular (`C01-C10`), analgesics (`N02`), and respiratory (`R01-R07`).

**Discussion:**
DPNs reveal key polypharmacy patterns. High-centrality medications highlight potential targets for drug monitoring. This approach aids in identifying influential medications and refining prescribing oversight.

>>>>

**Background:**
Drug prescription networks (DPNs) model the temporal dynamics of medication co-prescription within a population. Understanding prescription networks can provide insights into prescribing behaviors.

**Objectives:**
This study assesses the characteristics of temporal DPNs derived from daily co-prescriptions of antidepressants, anxiolytics, and all other therapeutic drug classes. By analyzing these networks using eigenvector centrality, we identified influential medications and prescribing patterns.

**Methods:**
Data were obtained from the University of Groningen IADB.nl database, including prescriptions from 128 Dutch pharmacies (2018–2022). A static cohort of patients prescribed antidepressants (ATC `N06A`) / anxiolytics (`N05B`) was extracted. Medications were classified using the Anatomical Therapeutic Chemical (ATC) system into 24 therapeutic classes. Time-varying DPNs were constructed as undirected graphs using symmetric daily dose-adjusted co-prescriptions. Eigenvector centrality ($c_e$) quantified relative nodal importance. Weekly-aggregated data included prescriptions ($n_c$), number of patients, dispensing-to-patient ratio, and eigenvector centrality. Singular spectrum analysis decomposed trends.

**Results:**
Antidepressants ($c_e$: `r N06A$ce`, $n_c$: `r N06A$nc`) and anxiolytics ($c_e$: `r N05B$ce`, $n_c$: `r N05B$nc`) had high eigenvector centrality, indicating frequent co-prescription. Additional high-centrality groups included medications for alimentary and metabolism (`A01-A16`), blood (`B01-B06`), cardiovascular (`C01-C10`), analgesics (`N02`), and respiratory diseases (`R01-R07`).

**Discussion:**
DPNs revealed key co-prescription patterns. High-centrality medications highlight potential targets for drug monitoring.

### 2025-05-07 docs/article.qmd
Focus the manuscript on co-prescription instead of polypharmacy

A DPN is a graphical representation where medications correspond to nodes, and edges represent co-prescription relationships between them. Typically, simple undirected graphs are represented by a square symmetric adjacency matrix $A$, where each nonzero element $a_{ij}$ indicates that medications $i$ and $j$ were co-prescribed on a given day. A population-level DPN can be constructed by aggregating individual patient networks through matrix addition [@Cavallo2012]. Analyzing the structural characteristics of a DPN using both local (medication-level) and global (network-level) graph measures provides insights into co-prescription dynamics.

Node centrality measures, such as eigenvector centrality, quantify the importance of a medication in polypharmacy patterns [@Askar2021]. Addressing concerns about standardized polypharmacy indicators raised by @Sirois2016 and @Delara2022, we highlight eigenvector centrality as a key measure of a medication’s connection to highly prescribed drugs. A high eigenvector centrality suggests frequent co-prescription with commonly used medications, reflecting prescribing trends. Building on prior studies [@Cavallo2012; @Bazzoni2015], we hypothesize that eigenvector centrality serves as a meaningful quantitative indicator of polypharmacy. Therefore, this study aims to assess the structural characteristics of temporal DPNs, focusing on the role of eigenvector centrality in identifying influential medications and prescribing patterns in mental health treatments.

@@

## Source of data

@@

A graph can be represented as an adjacency matrix $A$. The adjacency matrix $A$ has dimensions of $|\mathcal{V}| \times |\mathcal{V}|$, where $|\mathcal{V}|$ is the number of nodes in a graph. Each element of $a_{ij}$ is the edge, where in an unweighted graph, $a_{ij} = 1$ if a connection exists between $\mathcal{v}_i$ and $\mathcal{v}_j$, and $a_{ij} = 0$ otherwise. In a weighted graph, $a_{ij}$ can take any real value ($a_{ij} \in \mathbb{R}$) depending on the application [@estrada2012structure].

@@

The choice of centrality largely depends on the objective of network analysis. As a general guide, degree centrality is useful to identify popular medication and monitor drug overuse. Betweenness centrality is suitable for targeting for drug-interaction study and optimizing therapy plan. Closeness centrality indicates widely-used key medications and efficiency in treatment networks. Eigenvector centrality is helpful to identify influential medications and narrow down high-impact medications for drug monitoring. This study focused on eigenvector centrality to evaluate which medications have a significant influence on how antidepressants and anxiolytics are prescribed. @eq-eigen-centrality outlines the calculation of eigenvector centrality; $c_i$ and $c_j$ are the centralities of node $\mathcal{v}_i$ and $\mathcal{v}_j$, respectively; $\lambda$ is the eigenvalue; $a_{ji}$ is an element on row $j$ and column $i$ from an adjacency matrix $A$, representing the connection between node $\mathcal{v}_j$ and $\mathcal{v}_i$.

$$
\displaystyle c_i = \frac{1}{\lambda} \sum_{j \neq i} a_{ji} \cdot c_j
$$ {#eq-eigen-centrality}

@@

Since the sum of all eigenvector centralities $c_i$ in equation [-@eq-eigen-centrality] is equal to one, a network with similarly connected nodes will have an eigenvector centrality $c_i$ = $\frac{1}{n}$ for each node, where $n$ is the number of nodes. The value of $c_i = \frac{1}{n}$ is the expected eigenvector centrality in a densely connected network where the occurrence of an edge $\{\mathcal{v}_i, \mathcal{v}_j\} \in \mathcal{E}$ has the same probability for all nodes. Per the central limit theorem, the value $\frac{1}{n}$ can be parameterized as a mean of a normal distribution, where all instances of eigenvector centrality will follow the distribution, $c_i \sim \mathcal{N}(\mu, \sigma)$. A one-sample Student's T-test was performed to assess how much each eigenvector centrality differs from the expected values. As such, eigenvector centrality greater than $\frac{1}{n}$ can be considered high, and the value below it can be considered low. The p-values was corrected for a type-I error using Bonferroni method.

@@

## Weekly variation of medication dispensing

The highest average of medication dispensing occurred on `r hi_claim$day`, with an average of `r hi_mean` [SD: `r hi_claim$sd_claim`]. Meanwhile, the lowest average of medication dispensing occurred on `r lo_claim$day`, with an average of `r lo_mean` [SD: `r lo_claim$sd_claim`] dispensing. A large margin of difference between the highest and lowest average number of dispensing implies daily variation and cyclicality, where it peaks on Monday, starts declining afterwards, and is at the lowest during weekends, as shown in @tbl-overview-daily. The cyclical patterns are detailed in the supplementary material section S2.1.

@@

For brevity, @fig-ssa only depicts the decomposition result of antidepressants dispensing over the years. SSA-based decomposition separates the trend from its harmonics, as shown in the lower half of the plot. The trend explains roughly 90% of the variability, and harmonics explain 10%. The first two harmonics, F1 and F2, captured most of the variability in a series compared to the rest of it. The trend, F1, and F2 were used to reconstruct the time series. The largest graph in @fig-ssa represents the reconstructed time-series (green line) and its original data points (gray line). Other decomposition results are reported in the supplementary file section S2.2.

@@

Seven medication classes exhibited high centrality, as shown in @fig-hi-eigen and @tbl-desc. Notably, highly dispensed medications generally had high eigenvector centrality, which aligns with the theoretical expectations from @eq-eigen-centrality. However, despite a higher number of dispensing (`r diffMed(antidepressants, respiratory, reportMedClaim)`), antidepressants had lower eigenvector centrality than medications for the respiratory system (`r diffMed(antidepressants, respiratory, reportMedEigen)`). A similar pattern was observed for anxiolytics, which had more dispensing than analgesics (`r diffMed(anxiolytics, analgesics, reportMedClaim)`) but lower eigenvector centrality (`r diffMed(anxiolytics, analgesics, reportMedEigen)`).

@@

This network analysis of dispensing data in individuals using antidepressants or anxiolytics revealed distinct co-prescription patterns between these groups. While same-class co-prescription was more common among antidepressant users, multi-class co-prescription predominated among those receiving anxiolytics. Notably, in both groups, multi-class co-prescription was at least ten times more prevalent than same-class co-prescription. Seven medication classes exhibited high eigenvector centrality, indicating their strong influence on co-prescription dynamics. These include medications for the alimentary tract and metabolism (ATC code `A`), blood and blood-forming organs (`B`), cardiovascular system (`C`), respiratory system (`R`), and analgesics (`N02`). These classes do not merely appear frequently but also serve as key connectors within prescribing networks, linking psychiatric treatments to chronic disease management. Their centrality suggests that they play a role in structuring multi-class co-prescription, either as standard co-prescriptions due to clinical guidelines or as necessary adjuncts in managing comorbidities.

Our findings align with previous research showing that antidepressants and anxiolytics are frequently co-prescribed with other medications [@Shrivastava2013], and that multi-class regimens contribute most to psychopharmaca polypharmacy [@de2004polypharmacy]. Furthermore, there have been established bidirectional associations between depression/anxiety and chronic illnesses [@qi2024longitudinal]. Conditions such as diabetes mellitus, thyroid disorders, and asthma have been linked to an increased risk of depression [@jang2024temporal]. The need for long-term pharmacological management further drives co-prescription patterns. Additionally, chronic illness-related anxiety can contribute to heightened prescribing of anxiolytics [@lebel2020health]. This underscores the interconnected nature of psychopharmaca prescribing patterns and highlights the complexity of medication management in these populations.

@@

This study utilized DPN to evaluate the influence of specific medication classes on prescribing patterns of antidepressants and anxiolytics. By focusing on co-prescription patterns, we identified seven medication classes with significant eigenvector centrality, reflecting their critical role in polypharmacy regimens for individuals using anxiolytics and antidepressants. This study highlights the potential of network-based approaches to narrow down high-impact medications for further evaluation.

>>>>

A DPN is a graphical representation where medications correspond to nodes, and edges represent co-prescription relationships between them. Typically, simple undirected graphs are represented by a square symmetric adjacency matrix $A$, where each nonzero element $A_{ij}$ indicates that medications $i$ and $j$ were co-prescribed on a given day. A population-level DPN can be constructed by aggregating individual patient networks through matrix addition [@Cavallo2012]. Analyzing the structural characteristics of a DPN using both local (medication-level) and global (network-level) graph measures provides insights into co-prescription dynamics.

Node centrality measures, such as eigenvector centrality, quantify the importance of a medication in polypharmacy patterns [@Askar2021]. Addressing concerns about standardized polypharmacy indicators raised by @Sirois2016 and @Delara2022, we highlight eigenvector centrality as a key measure of a medication’s connection to highly prescribed drugs. A high eigenvector centrality suggests frequent co-prescription with commonly used medications, reflecting prescribing trends. Building on prior studies [@Cavallo2012; @Bazzoni2015], we hypothesize that eigenvector centrality serves as a meaningful quantitative indicator of polypharmacy. Therefore, this study aimed to assess the structural characteristics of temporal DPNs, focusing on the role of eigenvector centrality in identifying associated medications in adult patients prescribed for antidepressant or anxiolytic.

@@

## Source of data and study population

@@

A graph can be represented as an adjacency matrix $A$. The adjacency matrix $A$ has dimensions of $|\mathcal{V}| \times |\mathcal{V}|$, where $|\mathcal{V}|$ is the number of nodes in a graph. Each element of $A_{ij}$ is the edge, where in an unweighted graph, $A_{ij} = 1$ if a connection exists between $\mathcal{v}_i$ and $\mathcal{v}_j$, and $A_{ij} = 0$ otherwise. In a weighted graph, $A_{ij}$ can take any real value ($A_{ij} \in \mathbb{R}$) depending on the application [@estrada2012structure].

@@

The choice of centrality largely depends on the objective of network analysis. As a general guide, degree centrality is useful to identify popular medication and monitor drug overuse. Betweenness centrality is suitable for targeting for drug-interaction study and optimizing therapy plan. Closeness centrality indicates widely-used key medications and efficiency in treatment networks. Eigenvector centrality is helpful to identify influential medications and narrow down high-impact medications for drug monitoring. We defined high and low eigenvector centrality relative to the expected value $\frac{1}{n}$, based on the null model of uniform connectivity (see "Determining relative importance"). This study focused on eigenvector centrality to evaluate which medications have a significant influence on how antidepressants and anxiolytics are prescribed. @eq-eigen-centrality outlines the calculation of eigenvector centrality; $c_i$ and $c_j$ are the centralities of node $\mathcal{v}_i$ and $\mathcal{v}_j$, respectively; $\lambda$ is the eigenvalue; $A_{ji}$ is an element on row $j$ and column $i$ from an adjacency matrix $A$, representing the connection between node $\mathcal{v}_j$ and $\mathcal{v}_i$.

$$
\displaystyle c_i = \frac{1}{\lambda} \sum_{j \neq i} A_{ji} \cdot c_j
$$ {#eq-eigen-centrality}

@@

Since the sum of all eigenvector centralities $c_i$ in equation [-@eq-eigen-centrality] equals 1, the expected eigenvector centrality for each node in a uniformly connected network is $\frac{1}{n}$, where $n$ is the total number of nodes. In such a network, each node has an equal probability of being connected to any other, and thus no node is more "influential" than another. This expected value value, $\frac{1}{n}$, serves as a baseline under a null model of uniform connectivity. Assuming eigenvector centrality scores approximately follow a normal distribution, we applied a one-sample Student's T-test with Bonferroni correction to assess how much each eigenvector centrality differs from the expected values. Nodes with $c_i$ greater than $\frac{1}{n}$ were categorized as having high eigenvector centrality, and those below the expected value were considered low.

@@

## Cyclicality and seasonality of medication dispensing

Visual inspection of daily dispensing volumes revealed consistent weekly cycles. Dispensing peaked on `r hi_claim$day`, with an average of `r hi_mean` [SD: `r hi_claim$sd_claim`] and dropped to its lowest on `r lo_claim$day`, with an average of `r lo_mean` [SD: `r lo_claim$sd_claim`] dispensing. The variation between weekends and weekdays suggested a regular and repeating structure in the data. A large margin of difference between the highest and lowest average number of dispensing implied daily variation and cyclicality, where it peaks on Monday, starts declining afterwards, and is at the lowest during weekends, as shown in @tbl-overview-daily. The cyclical patterns are detailed in the supplementary material section S2.1. Seasonality was explored as dependencies at daily and weekly lags and further described in the supplementary material section S2.2.

@@

For brevity, @fig-ssa only depicts the decomposition result of antidepressants dispensing over the years. SSA-based decomposition separates the trend from its harmonics, as shown in the lower half of the plot. The trend explains roughly 90% of the variability, and harmonics explain 10%. The first two harmonics, F1 and F2, captured most of the variability in a series compared to the rest of it. The trend, F1, and F2 were used to reconstruct the time series. The largest graph in @fig-ssa represents the reconstructed time-series (green line) and its original data points (gray line). Other decomposition results are reported in the supplementary file section S2.3.

@@

Seven medication classes exhibited high centrality, as shown in @fig-hi-eigen and @tbl-desc. Notably, highly dispensed medications generally had high eigenvector centrality, consistent with the theoretical expectations from @eq-eigen-centrality. However, this relationship was not strictly proportional. For example, although antidepressants had a higher number of dispensing compared to medications for the respiratory system (`r diffMed(antidepressants, respiratory, reportMedClaim)`), they exhibited lower eigenvector centrality (`r diffMed(antidepressants, respiratory, reportMedEigen)`). A similar pattern was observed for anxiolytics, which had more dispensing than analgesics (`r diffMed(anxiolytics, analgesics, reportMedClaim)`) but lower eigenvector centrality (`r diffMed(anxiolytics, analgesics, reportMedEigen)`). This reflects that eigenvector centrality captures not just volume but also the influence of connection sin the network, highlighting that commonly dispensed drug within a population may not necessarily occupy the most central network positions.

@@

This network analysis of dispensing data in individuals using antidepressants or anxiolytics revealed distinct co-prescription patterns. While same-class co-prescription was more common among antidepressant users, multi-class co-prescription predominated among those receiving anxiolytics. Notably, in both groups, multi-class co-prescription was at least ten times more prevalent than same-class co-prescription. Seven medication classes exhibited high eigenvector centrality, indicating their strong influence on co-prescription dynamics. These include medications for the alimentary tract and metabolism (ATC code `A`), blood and blood-forming organs (`B`), cardiovascular system (`C`), respiratory system (`R`), and analgesics (`N02`). These classes do not merely appear frequently but also serve as key connectors within prescribing networks, linking psychiatric treatments to chronic disease management. Their centrality suggests that they play a role in structuring multi-class co-prescription, either as standard co-prescriptions due to clinical guidelines or as necessary adjuncts in managing comorbidities or side effects, e.g., ATC `A` class.

The predomination of multi-class co-prescription among anxiolytics recipients is consistent with the recommendations in the Dutch Huisartsengenootschap [NHG; @NHG2019Anxiety]. Specifically, the NHG guideline for anxiety disorders advises short-term benzodiazepine use, limited to two to four weeks, as an adjunct to Selective Serotonin Reuptake Inhibitors (SSRIs) during treatment initiation. In contrast, the NHG guideline for depressive disorders recommends initiating pharmacotherapy with a single SSRI and discourages the combination of multiple antidepressants [@NHG2023Depression]. Therefore, the higher rate of multi-class co-prescription observed among anxiolytic users is consistent with evidence-based clinical practice, rather than an artifact of sampling or measurement methods.

Our findings also align with previous research showing that antidepressants and anxiolytics are frequently co-prescribed with other medications [@Shrivastava2013], and that multi-class regimens contribute most to psychopharmaca polypharmacy [@de2004polypharmacy]. Additionally, our study offers a more granular view of these patterns by characterizing their network properties. By distinguishing between same-class and multi-class co-prescriptions and evaluating medication centrality, we discover that certain drug classes function as hubs within the prescribing landscape. These characteristics likely reflect the underlying burden of multimorbidity in the population, where the bidirectional associations between depression/anxiety and chronic illnesses is well established [@qi2024longitudinal]. Conditions such as diabetes mellitus, thyroid disorders, and asthma have been linked to an increased risk of depression [@jang2024temporal]. The need for long-term pharmacological management further drives co-prescription patterns. Additionally, chronic illness-related anxiety can contribute to heightened prescribing of anxiolytics [@lebel2020health]. This underscores the interconnected nature of psychopharmaca prescribing patterns and highlights the complexity of medication management in these populations.

@@

This study demonstrates the utility of DPN in uncovering the patterns of co-prescriptions among antidepressant and anxiolytic recipients. We found that multi-class regimens were markedly more prevalent, particularly among anxiolytic users, consistent with clinical guidelines. Moreover, we identified seven highly central medication classes that act as hubs in the prescribing network, linking psychiatric medications to the management of chronic somatic conditions. These findings emphasize the role of multimorbidity in shaping co-prescription behaviors and highlight the importance of these seven classes in both psychiatric care and broader chronic disease management. Our study contributes a network-based perspective to drug utilization research, offering an alternative approach to identifying high-impact medication classes that warrant further monitoring and future investigation.

### 2025-05-07 docs/exec-summary/_preprocessing.qmd
Explain table 2 in the result section
Focus the manuscript on co-prescription instead of polypharmacy

Daily medication claims are represented as tabular data where each row corresponds to an individual patient’s prescriptions. The data was transformed into a square matrix summarizing co-prescriptions and single prescriptions to construct a graph-based representation. The rows and columns of the matrix represent groups of medications, while each matrix element indicates the frequency of a specific prescription pattern. Elements in the off-diagonal positions represent co-prescriptions between two medications, while diagonal elements represent single prescriptions. This matrix can be interpreted as a graph, where each medication is a node, and the edges represent co-prescription frequencies. The weight of the edge is derived from the number of daily claims. Originally, all defined daily doses (DDD) were assumed to equal 1, leading to straightforward edge weights based on raw counts. However, since not all medications have DDD values of 1, this assumption could lead to inaccurate representations. To address this, weights were adjusted using the density of a normal distribution. The expected value ($\mu$) was set to 1 to prioritize DDD = 1, and the standard deviation ($\sigma$) was set to $\frac{1}{3}$ to allow for a gradual reduction in weight as DDD deviates from 1. This adjustment ensures a more accurate weighting system that reflects real-world prescription patterns while preserving information about medications with non-standard DDD values.

>>>>

At the individual level, each patient's daily prescription record at time $t$ was encoded into a square matrix $A_t^{(i)}$, where rows and columns represent medication classes according to the ATC classification. An off-diagonal entry $A_{t_{jk}}^{(i)} = 1$ indicates that medication classes $i$ and $k$ were co-prescribed on the same day for individual $i$, while a diagonal entry $A_{t_{jj}}^{(i)} = 1$ captures the presence of a single prescription of medication $j$. To derive the population-level DPN, all individual-level daily prescription matrix $A^{(i)}$ were aggregated by computing their element-wise sum.

$$A_t = \displaystyle \sum_i A_t^{(i)}$$ {#eq-DPN}

The population matrix $A_t$ captures the frequency of co-prescriptions (off-diagonal) and single prescriptions (diagonal) across the dataset at the specified time $t$. Each element $A_{t_{jk}}$ in the matrix was interpreted as the number of patients being prescribed with medication classes $j$ and $k$ on the same day $t$. $A_t$ was represented as a weighted undirected graph, where each node corresponds to a medication class and edges reflect co-prescription frequencies. In the original work of DPN, all defined daily dose (DDD) were assumed equal to 1, allowing for applying raw counts as the edge weights. However, since not all medications have DDD = 1, this assumption could lead to inaccurate representations. To enhance the accuracy of edge weights, we accounted for the heterogeneity in DDD. The weights were adjusted using the density of a normal distribution. The expected value ($\mu$) was set to the base weight $\omega_B$ of 1 to prioritize DDD = 1, and the standard deviation ($\sigma$) was set to $\frac{1}{3}$ to allow for a gradual reduction in weight as DDD deviates from 1.

### 2025-05-07 docs/exec-summary/_source-data.qmd
Focus the manuscript on co-prescription instead of polypharmacy

All data in this investigation originated from IADB.nl, a dynamic pharmacy database containing prescription data from 1994 which still grow expansively. The database covers approximately 120 community pharmacies serving over one million patients. All patients are recorded in the database, irrespective of health care insurance, where the prescription rates, age, and gender are generalizable to represent the Netherlands [@Visser2013]. The database offers a longitudinal record, where each individual has a uniquely anonymized tracking identifier. All records contain fields on the date of medication dispensing, its quantity, dose regimen, number of days prescribed, prescribing physicians, and the Anatomical Therapeutic Chemical (ATC) code. Except for the over the counter drugs and medication dispensed during hospitalization, other medication records for each patient are complete, mainly due to high patient-pharmacy commitment in the Netherlands. This analysis include daily medication claims from a static cohort of adults aged 18-65 years old and prescribed for anxiolytics or antidepressants at least once in the time span from 2018 to 2022.

>>>>

All data in this study originated from the University of Groningen IADB.nl, a dynamic pharmacy database containing prescription data from 1994 which still grows expansively. The database covers approximately 128 community pharmacies serving over one million patients and was accessed on the 20^th^of September, 2023. All patients are recorded in the database, irrespective of health care insurance, where the prescription rates, age, and gender are generalizable to represent the Netherlands [@Visser2013]. The database offers a longitudinal record, where each individual has a uniquely anonymized tracking identifier. All records contain fields on the date of medication dispensing, its quantity, dose regimen, number of days prescribed, prescribing physicians, and the Anatomical Therapeutic Chemical (ATC) code. Except for the over the counter drugs and medication dispensed during hospitalization, other medication records for each patient are complete, mainly due to high patient-pharmacy commitment in the Netherlands.

This analysis includes daily medication dispensing from a static cohort of adults aged 18-65 years old and prescribed for anxiolytic or antidepressant at least once in the time span from 2018 to 2022.

### 2025-05-07 docs/supplementary.qmd
Focus the manuscript on co-prescription instead of polypharmacy

```{r fig-period-day}
#| fig-width: 14
#| fig-height: 5
#| fig-cap: Daily cyclical pattern captured on a de-trended data

@@

@@

## Time-series decomposition

```{r}
#| fig-width: 16
#| fig-height: 12
#| fig-pos: H
#| fig-cap: Time-series decomposition using singular spectrum analysis-based approach

>>>>

```{r fig-period-day}
#| fig-width: 14
#| fig-height: 8
#| fig-cap: Daily cyclical pattern captured on a de-trended data

@@

\pagebreak

## Daily ACF and PACF plots

```{r fig-season-day}
#| fig-width: 14
#| fig-height: 8
#| fig-cap: Daily cyclical pattern captured on a de-trended data

tar_read(plt_acf_day)[[4]]
tar_read(plt_pacf_day)[[4]]

\pagebreak

## Weekly ACF and PACF plots

```{r fig-season-week}
#| fig-width: 14
#| fig-height: 8
#| fig-cap: Daily cyclical pattern captured on a de-trended data

tar_read(plt_acf_week)[[4]]
tar_read(plt_pacf_week)[[4]]

@@

## Time-series decomposition of medication dispensing records

```{r}
#| fig-width: 16
#| fig-height: 9
#| fig-pos: H
#| fig-cap: Time-series decomposition using singular spectrum analysis-based approach

### 2025-05-07 docs/supplementary.tex
Focus the manuscript on co-prescription instead of polypharmacy

In case of complex periodic patterns, sequential SSA is preferred to
separate trend from seasonal components (Golyandina, Korobeynikov, and
Zhigljavsky 2018). A sequential SSA model is performed by first fitting
a basic SSA model to reconstruct the first eigentriple, \emph{viz.} the
eigentriple that explain most of the variance in the data. Then, the
residual of the first model is extracted to fit the second SSA model.
The first SSA model is defined with the lowest possible \(L\), which in
our case \(L = 52\). The second SSA model is defined with the largest
possible \(L\) to capture all patterns, which is conveniently defined as
\(L = \frac{N}{2}\). The decomposed series and its corresponding
original data is then extracted from the model alongside all of the
oscillating functions which contributes to complex seasonality. The
original data, trend, residuals, and oscillating functions are plotted
for further inspection. The original data and trend is further evaluated
using Mann-Kendall trend test.

@@

@@

\subsection{Time-series decomposition}\label{time-series-decomposition}

>>>>

In case of complex periodic patterns, sequential SSA is preferred to
separate trend from seasonal components (Golyandina, Korobeynikov, and
Zhigljavsky 2018). A sequential SSA model is performed by first fitting
a basic SSA model to reconstruct the first eigentriple, \emph{viz.} the
eigentriple that explain most of the variance in the data. Then, the
residual of the first model is extracted to fit the second SSA model.
The first SSA model is defined with the lowest possible \(L\), which in
our case \(L = 52\). The second SSA model is defined with the largest
possible \(L\) to capture all patterns, which is conveniently defined as
\(L = \frac{N}{2}\). The decomposed series and its corresponding
original data is then extracted from the model alongside all of the
oscillating functions which contributes to complex seasonality. The
oscillatory components extracted from SSA correspond to harmonics,
namely a distinct periodic signals with different frequencies. These
harmonics form complex seasonality and captured in pairs of
eigentriples. The original data, trend, residuals, and oscillating
functions are plotted for further inspection. The original data and
trend is further evaluated using Mann-Kendall trend test.

@@

\pagebreak

\subsection{Daily ACF and PACF plots}\label{daily-acf-and-pacf-plots}

\includegraphics[width=1\linewidth,height=\textheight,keepaspectratio]{supplementary_files/figure-pdf/fig-season-day-1.pdf}

\caption{\label{fig-season-day-1}Daily cyclical pattern captured on a
de-trended data}

\includegraphics[width=1\linewidth,height=\textheight,keepaspectratio]{supplementary_files/figure-pdf/fig-season-day-2.pdf}

\caption{\label{fig-season-day-2}Daily cyclical pattern captured on a
de-trended data}

\pagebreak

\subsection{Weekly ACF and PACF plots}\label{weekly-acf-and-pacf-plots}

\includegraphics[width=1\linewidth,height=\textheight,keepaspectratio]{supplementary_files/figure-pdf/fig-season-week-1.pdf}

\caption{\label{fig-season-week-1}Daily cyclical pattern captured on a
de-trended data}

\includegraphics[width=1\linewidth,height=\textheight,keepaspectratio]{supplementary_files/figure-pdf/fig-season-week-2.pdf}

\caption{\label{fig-season-week-2}Daily cyclical pattern captured on a
de-trended data}

@@

\subsection{Time-series decomposition of medication dispensing
records}\label{time-series-decomposition-of-medication-dispensing-records}
